Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-11-15 paper/main.tex
Added modification of tahoe code sec

@@

\paragraph{Security.} The most important criteria for the SHA-3
candidates\cite{s_nistround2} is security. The full description of security
criteria are mentioned in \cite{s_fedreg}, however the most noteworthy criterion
in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
family are vulnerable to this kind of attack, and thus requires
Tahoe to run the algorithm twice, as described by \citet{schneier}.

@@

In this chapter we will outline the procedure followed to complete our task.

@@

\section{Modifications in the Tahoe-LAFS Code}

% This should be a short text, but maybe drag in a discussion of how we SHOULD
% have implemented it if we had the time, ref. capabilities.

@@

\subsection{Physical Setup}

>>>>

% Packages and settings for code listings
\usepackage{listings}
\usepackage{caption}
\usepackage{upquote}
\usepackage{xcolor}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
\lstset{
language=Python,
keywordstyle=\bfseries\ttfamily\color[rgb]{0,0,1},
identifierstyle=\ttfamily,
commentstyle=\color[rgb]{0.133,0.545,0.133},
stringstyle=\ttfamily\color[rgb]{0.627,0.126,0.941},
showstringspaces=false,
basicstyle=\small,
numberstyle=\footnotesize,
numbers=left,
stepnumber=1,
numbersep=10pt,
tabsize=2,
breaklines=true,
prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
breakatwhitespace=false,
aboveskip={1.5\baselineskip},
columns=fixed,
upquote=true,
extendedchars=true,
frame=bottomline,
inputencoding=utf8
}

@@

\label{sec:lengthextension}
\paragraph{Security.} The most important criteria for the SHA-3
candidates\cite{s_nistround2} is security. The full description of security
criteria are mentioned in \cite{s_fedreg}, however the most noteworthy criterion
in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
family are vulnerable to this kind of attack, and thus requires
Tahoe to run the algorithm twice, as described by \citet{schneier}.

@@

In this chapter we will outline the procedure followed to complete our task.
% TODO: Mer intro

@@

\section{Modifications in the Tahoe-\ac{LAFS} Code}

The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
This, along with the code being highly modular, emphasizes code readability.
With some experience using Python, it is not difficult to learn the Tahoe code
enough to understand what is happening when a file is uploaded or downloaded
with the system.

All hashing operations requested by the Tahoe system is routed through a
couple of classes and functions, and the connection to the underlying hash
function library is made on only one line. This is displayed in Listing
\ref{lst:hashutil}.

\begin{lstlisting}[label=lst:hashutil, caption=Extract from hashutil.py of Tahoe-LAFS source.]
from pycryptopp.hash.sha256 import SHA256
...
class _SHA256d_Hasher:
    ...
    def __init__(self, truncate_to=None):
        self.h = SHA256()
        self.truncate_to = truncate_to
        self._digest = None

    def update(self, data):
        assert isinstance(data, str) # no unicode
        self.h.update(data)

    def digest(self):
        if self._digest is None:
            h1 = self.h.digest()
            del self.h
            h2 = SHA256(h1).digest()
            if self.truncate_to:
                h1 = h1[:self.truncate_to]
                h2 = h2[:self.truncate_to]
            self._digest = h2

        return self._digest
\end{lstlisting}

As stated in Section \ref{sec:lengthextension}, none of the SHA-3 candidates is
vulnerable to the length-extension attack. Since we are interested in finding
out how Tahoe-\ac{LAFS} performs when utilizing the SHA-3 functions, we can
remove the lines that fixes the length-extension problems manually. In addition,
to try out one of the new implementations, we change line one in Listing
\ref{lst:hashutil}, and the result can be seen in Listing \ref{lst:hashutilmod}.

\begin{lstlisting}[label=lst:hashutilmod, caption=Relevant parts of hashutil.py of Tahoe-LAFS source after modification.]
from sha3lib import bmw256 as SHA256
...
class _SHA256d_Hasher:
    ...
    def __init__(self, truncate_to=None):
        self.h = SHA256()
        self.truncate_to = truncate_to
        self._digest = None

    def update(self, data):
        assert isinstance(data, str) # no unicode
        self.h.update(data)

    def digest(self):
        if self._digest is None:
            h1 = self.h.digest()
            del self.h
            if self.truncate_to:
                h1 = h1[:self.truncate_to]
            self._digest = h1

        return self._digest
\end{lstlisting}

Some notes has to be made regarding the first line in Listing
\ref{lst:hashutilmod}. For every candidate we test, the whole setup and grid is
started from ground up, since changing the hash function this way cause a
backwards compatibility break, making it impossible to download files put to the
grid while using another hash function.
For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with SHA-256.
This also renders uploading of mutable files impossible, since the internal
verification fails. For this reason, \emph{only immutable files} are tested and
measured in this paper.

@@

% TODO: Write something about why we only test the client
